VOI-aware Monte-Carlo Sampling in Trees

### 2011-12-28 paper/vago/paper.tex
refinement to the paper

For the experiments, a modified version of Pachi \cite{Braudis.pachi},
a state of the art Go program, was used:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping condition for the VOI-aware policies was
  modified and based solely on the sample cost, specified as
  a constant parameter. 
\item The time-allocation mode based on the fixed number of samples
  was modified such that 
  \begin{enumerate}
    \item the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item if samples were unused at the current step,
      they become available at the next step.
  \end{enumerate}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still
a strong player. On the other hand, additional
features of more advanced engines would obstruct the Monte-Carlo
sampling phenomena which are the subject of the experiment.

@@

As the results show, without sample redistribution UCT is as good as or
better than the VOI-aware policies for lower numbers of samples per ply
(5000, 7000), and the advantage of VOI-aware policies for larger
number of samples is rather modest (e.g., ECT wins in 55\% of games for
10000 samples per ply). For the stopping threshold of $10^{-6}$,
however, the VOI-aware policies are almost always better than UCT
(only VCT for 5000 samples wins in less than 50\% of games), with
stronger policies (ECT, VCT) reaching the winning rate of 66\%.

>>>>

For the experiments, a modified version of Pachi \cite{Braudis.pachi},
a state of the art Go program, was used:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping condition for the VOI-aware policies was
  modified and based solely on the sample cost, specified as
  a constant parameter. 
\item The time-allocation mode based on the fixed number of samples
  was modified for both the original UCT policy and the VOI-aware
  policies such that 
  \begin{enumerate}
    \item the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item if samples were unused at the current step,
      they become available at the next step. 
  \end{enumerate}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is
still a strong player. On the other hand, additional features of more
advanced engines would obstruct the Monte-Carlo sampling phenomena
which are the subject of the experiment.

@@

As the results show, without VOI-aware sample redistribution the
VOI-aware policies are at best equal to UCT for lower numbers of
samples per ply (5000, 7000), and the advantage of the VOI-aware policies
for larger number of samples is rather modest (e.g., ECT wins in 55\%
of games for 10000 samples per ply). For the stopping threshold of
$10^{-6}$, however, the VOI-aware policies are almost always better
than UCT (only VCT for 5000 samples wins in less than 50\% of games),
with stronger policies (ECT, VCT) reaching the winning rate of 66\%.
